---
title: "Culminating Activity"
author: "Alagon, Arcena, Argawanon, Coyacot, Dalocanog"
date: "2026-05-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(dplyr)
library(caret)
library(class)
```

# Abstract
Provide a concise summary of the study including purpose, methods, results, and conclusions (150-250 words)

# Chapter 1: Introduction
## Background of the Study
Discuss the context and importance of the research problem

## Statement of the Problem
Clearly state the research questions.

1. What is... ?

2. How does... ?

## Objectives of the Study
## General Objective

- To investigate ...

## Specific Objectives

- To determine ...

- To analyze ...

## Hypotheses (if applicable)

- **H0:** There is no significant relationship between...

- **H1:** There is a significant relationship between...

## Significance of the study
Explain who benefits and how.

## Scope and Limitations
Define the boundaries of the study.

# Chapter 2: Review of Related Literature
## Related Studies
Summarize previous research

## Related Literature
Discuss theories and concepts.

## Conceptual Framework
(Insert diagram if needed)

# Chapter 3: Methodology
Research Design

Describe the design (e.g., descriptive, experimental).

Data Collection

Get a data set from www.kaggle.com
```{r}
data <- read.csv('C:/Data Analysis/DataAnalysisCulminatingActivity/cs_students.csv')
kable(head(data),
      caption = "Descriptive Statistics of CS Students",
      align = "c")

data <- data[, !(names(data) %in% c("Student.ID", "Name"))]

data <- na.omit(data)

# Target variable
# data$Future.Career = factor(data$Future.Career)

# Weak = 1
# Average = 2
# Strong = 3
data$Python <- as.numeric(factor(data$Python,
                                 levels = c("Weak", "Average", "Strong")))
data$SQL <- as.numeric(factor(data$SQL,
                              levels = c("Weak", "Average", "Strong")))
data$Java <- as.numeric(factor(data$Java,
                              levels = c("Weak", "Average", "Strong")))

# Gender
data$Gender <- ifelse(data$Gender == "Male", 0, 1)

# Grouping 
library(dplyr)

data$Career_Group <- case_when(
  # Software Development
  data$Future.Career %in% c(
    "Software Engineer", "Web Developer", "Mobile App Developer",
    "Game Developer", "Embedded Software Engineer",
    "Distributed Systems Engineer", "Graphics Programmer"
  ) ~ "Software Dev",

  # Data & Analytics
  data$Future.Career %in% c(
    "Data Scientist", "Data Analyst", "Database Administrator",
    "Bioinformatician", "Geospatial Analyst"
  ) ~ "Data",

  # AI / Machine Learning
  data$Future.Career %in% c(
    "Machine Learning Engineer", "AI Researcher",
    "Machine Learning Researcher", "NLP Research Scientist",
    "NLP Engineer", "Computer Vision Engineer",
    "Robotics Engineer", "Quantum Computing Researcher"
  ) ~ "AI/ML",

  # Cybersecurity
  data$Future.Career %in% c(
    "Information Security Analyst", "Ethical Hacker",
    "Security Analyst", "Digital Forensics Specialist",
    "Data Privacy Specialist"
  ) ~ "Cybersecurity",

  # Cloud / Infrastructure
  data$Future.Career %in% c(
    "Cloud Solutions Architect", "DevOps Engineer",
    "IoT Developer"
  ) ~ "Cloud/DevOps",

  # Design & User Experience
  data$Future.Career %in% c(
    "UX Designer", "VR Developer"
  ) ~ "Design",

  # Business / Misc Tech
  data$Future.Career %in% c(
    "SEO Specialist", "Healthcare IT Specialist",
    "Blockchain Engineer"
  ) ~ "Other",

  TRUE ~ "Other"
)

data$Career_Group <- as.character(data$Career_Group)
data$Career_Group[is.na(data$Career_Group)] <- "Other"
data$Career_Group <- as.factor(data$Career_Group)

target <- data$Career_Group
features <- data[, !(names(data) %in% c("Future.Career", "Career_Group"))]

features <- features[, sapply(features, function(x) length(unique(x)) > 1)]

features_encoded <- model.matrix(~ . - 1, data = features)

sum(is.na(features_encoded))

# Change this to only calculate scaling parameters using the training data set
# This is to prevent data leakage from happening
# features_scaled <- scale(features_encoded)

target <- as.factor(target)

train_index <- createDataPartition(target,
                                   p = 0.7, 
                                   list = FALSE)

# train_index <- sample(1:nrow(features_encoded), 0.7 * nrow(features_encoded))

# train <- data[train_index, ]
# test <- data[-train_index, ]

features_encoded <- scale(features_encoded)

x_train <- features_encoded[train_index, ]
x_test  <- features_encoded[-train_index, ]

y_train <- target[train_index]
y_test  <- target[-train_index]

# sc <- scale(x_train)

# x_train <- sc
# x_test <- scale(x_test,
#                 center = attr(sc, "scaled:center"),
#                 scale  = attr(sc, "scaled:scale"))

# nzv <- apply(x_train, 2, sd) != 0
# x_train <- x_train[, nzv]
# x_test <- x_test[, nzv]

k_new <- round(sqrt(nrow(x_train)))
if (k_new %% 2 == 0) k_new <- k_new + 1

pred <- knn(train = x_train,
            test = x_test,
            cl = y_train,
            k = k_new)

pred

mean(pred == y_test)

# Confusion Matrix
cM = confusionMatrix(pred, y_test)
addmargins(cM$table)
```

Variables Description

Explain dependent and independent variables

Data Analysis Techniques

- Descriptive Statistics
- Inferential Statistics
- Machine Learning Models (if applicable)

# Chapter 4: Results and Discussion
Descriptive Statistics

Visualization

Model (Example: Linear Regression)

Interpretation

Discuss findings in paragraph form

# Chapter 5: Conclusion and Recommendations
Conclusion

Summarize key findings

**Recommendations**

Provide practical suggestions

## References
Use APA format.

## Appendices
Appendix A: Code